---
title: "Advanced Data Visualization in R"
author: Iris Malone
date: \today
output:
  beamer_presentation:
    theme: Madrid
---

```{r, echo=FALSE, results='hide'}
# Simple Example: Look at Waves of Democratization from 1800-2012
suppressMessages(require(psData))
df = suppressMessages(PolityGet(vars = 'polity2', OutCountryID = 'imf'))
df$dem = ifelse(df$polity2 >= 6, 1, 0)
df$aut = ifelse(df$polity2 <= -6, 1, 0)

# re-shape the data
year = unique(df$year)
numdems = tapply(df$dem, df$year, sum, na.rm = T)
numauts = tapply(df$aut, df$year, sum, na.rm = T)
```


# Outline
What I'm Covering: Good stuff. 

- pro/cons different graphing packages 
- ggplot and the grammar of graphics 
- visualizing summary stats, regression results 
- ggmap and basic map visualization 

What I'm Not Covering:  

- 3D Visualization. 
- Network Analysis, e.g. igraph 
- Interactive Visualization, e.g. Shiny, D3  \alert{Next VAM with Simon - Dec 4th!}

# Choosing a Visualization Package

Toy Example. 

Options:

- plot (graphics)

# Method 1: Plot
Code: 
```{r, results=FALSE, fig.show='hide'}
plot(year, numdems, #x, y 		
  #aesthetics stuff (color, plot type, size)
	col = "blue", type = "l" , lwd=3, 
  #main title and axes labels
	main = "Number of Democracies and Autocracies, 1800-2012", 
	xlab = "Year", 
	ylab = "Count")
lines(year, numauts, type = "l", col = "red", lwd =3) 
legend("topleft", #location of legend 
  legend=c("Democracies","Autocracies"), #legend labels/var
  col = c(4, 2), # colors for each, 
  #legend colors default 1= black, 2 = red, 4=blue
  lty = c(1, 1),
  title="Regime Type") # Name of Legend

```

#Option 1: Plot
Visual:

```{r, echo=FALSE}
plot(year, numdems, #x, y 		
  #aesthetics (color, plot type, size)
	col = "blue", type = "l" , lwd=3, 
  #main title and axes labels
	main = "Number of Democracies and Autocracies, 1800-2012", 
	xlab = "Year", 
	ylab = "Count")
lines(year, numauts, type = "l", col = "red", lwd =3) 
legend("topleft", #location of legend 
  legend=c("Democracies","Autocracies"), #legend labels/var
  col = c(4, 2), # colors for each, 
  #legend colors default 1= black, 2 = red, 4=blue
  lty = c(1, 1),
  title="Regime Type") # Name of Legend
```

# Choosing a Visualization Package

Toy Example. 

Options:

- plot (graphics)  \alert{Ancient.}  \alert{Annoying syntax.}  \alert{Not great support.} 
- xyplot (lattice)

# Option 2: xyplot

Step 1: Build First Layer for Number of Democracies.

```{r}
suppressMessages(require(lattice))
layer1 = xyplot(numdems ~ year, #format: y ~ x
	type = "l",
	# add title
	main = "Number of Democracies and Autocracies, 1800-2012", 
	col = "blue", lwd=1,
	#legend 
	key=list(space="right", # location 
	# aesthetics (aes)
  lines=list(col=c("red","blue"), lty=c(1,1), lwd=2), 
	#labels for each line
  text=list(c("Autocracies","Democracies")))) 
```

# Option 2: xyplot

Step 2: Build Second Layer for Number of Autocracies.

```{r}
layer2 = xyplot(numauts ~ year, 
		type = "l", 
		col = "red")
```

# Option 2: xyplot

Step 3: Add the layers.

```{r, fig.width=12, fig.height = 5}
#need extra package to put layers on top
suppressMessages(require(latticeExtra)) 
layer1 + layer2
```

# Graphics Packages
Options:

- plot (graphics)  \alert{Ancient.} \alert{Annoying syntax.} \alert{Not great support.} 
- xyplot (lattice)  \alert{How many layers?}  \alert{Weird rigidity.}  \alert{Ugly default style.} 
- ggplot (ggplot2)

# Option 3:

```{r, fig.show='hide'}
suppressMessages(require(ggplot2))
ggplot(data = NULL) + 
 geom_line(aes(x=year, y = numdems, colour = "numdems")) + #line for numdems
 geom_line(aes(x=year, y = numauts, colour = "numauts")) + #line for numauts
 xlab("Year") + ylab("Count") + 
 ggtitle("Number of Democracies and Autocracies, 1800-2012") + 	#title
 # legend aesthetics
 scale_color_manual(name = "Regime Type", # Name
 labels = c(numdems="Democracies", numauts = "Autocracies"), #labels for each var
	values=c(numdems=4,numauts=2))
```

# Option 3:

```{r, echo=FALSE, fig.width=12, fig.height = 6}

ggplot(data = NULL, aes(x = year)) + 
	geom_line(aes(y = numdems, colour = "numdems"), lwd = 1) + #line for numdems
	geom_line(aes(y = numauts, colour = "numauts"), lwd = 1) + #line for numauts
	xlab("Year") + # x-axis label
	ylab("Count") + #y-axis label
	ggtitle("Number of Democracies and Autocracies, 1800-2012") + 	#title
	scale_color_manual(name = "Regime Type", # Name of Legend
	labels = c(numdems="Democracies", numauts = "Autocracies"), #legend labels for each var
	values=c(numdems=4,numauts=2))
```

#Summary of Options:

- plot (graphics)
- xyplot (lattice)
- ggplot (ggplot2)

# Why ggplot?
- Very pretty  
- Used professionally 

# Why ggplot?
NYT.
\begin{center}
\includegraphics[width=11cm]{/users/irismalone/Dropbox/VAM/pics/nytggplot.png}
\end{center}

# Why ggplot?
\begin{center}
\includegraphics[width=8cm]{/users/irismalone/Dropbox/VAM/pics/nyt2.png}
\end{center}

# Why ggplot?
\begin{tabular}{rl}  
           & \begin{tabular}{r}
             \parbox{0.35\linewidth}{%  change the parbox width as appropiate
             - Very pretty \\
             - Used professionally \\
- Easy to manipulate  \\
- Great support online  \\
- Knowledge transfers to other packages/languages (ggvis, Shiny, Python)  \\
- \alert{Steep Learning Curve}  \\
- \alert{Lots of syntax}  \\
- \alert{Defaults to weird colors.}  \\

- \textbf{Summary: Worth it.}
    } \\
         \end{tabular}  
          \begin{tabular}{l}
        \includegraphics[width=5cm]{/users/irismalone/Dropbox/VAM/pics/dogedog.jpg}
           \end{tabular}
\end{tabular}

#Visualizing Basics

#Visualizing Regressions

# Visualizing Maps

# Summary
\begin{center}
\includegraphics[width=5cm]{/users/irismalone/Dropbox/VAM/pics/ggplot.jpg}
\end{center}

# Spatial Visualization

# ggmap


